Ionization of biological molecules by multicharged ions using the stoichiometric model

### 2019-12-02 submit_jpb/reviewers/response.txt

Detailed reply:

1) We have changed the wording of our findings in the corresponding sections.

2) The grammatical error was corrected.

3a) The dependence of the mean energy with the projectile charge at intermediate energy reflects the non-linear dependence on E and ZP of the continuum state in the CDW approximation. Then, the dependence of the numerator and the denominator on ZP is not canceled out, which explains the maximum in Fig. 2. On the contrary, the behavior at high energy follows the Z^2 rule, and the mean energy distribution tends to a universal value for all ions, as noted on page 3.
3b) We included a few sentences discussing the usefulness of understanding the relationship between the mean energy and angle distribution of secondary electrons with the radiation damage. 
3) We included a reference on radiation damage in the introduction as well.

4) The denominator in Eq. 4 is the same as in Eq. 3: the total cross-section. However, they are both integrated on different variables. We have included the corresponding expression of the normalization of the angular distribution.

@@

5) We defined the dotted line accordingly.

6a) We included the reference to Table I in the caption.
6b) In general, the curves within each scaling band are ordered from the smallest to the largest. For the hydrocarbons, this behavior holds for both scalings since n_e for carbon does not change. However, this is not always the case (see, for example, adenine C5H5N5, uracil C4H4N2O2, and DNA backbone C5H10O5P in figure A attached). 
6c) Upon the suggestion of the referee, we included the Toburen numbers in Table II. The inclusion of the band by Toburen, together with the new scaling, was too confusing and finally decided not to modify the figure.

7) We included the Toburen scaling in figure B, attached. The x-axis refers to the number of active electrons, whether they are from the Toburen or the CDW scaling. As seen, the scaling by Toburen is not as good as ours, which follows Fig. 6. To avoid having too many curves, we decided not to modify Fig. 7, but we comment on these differences in the manuscript: 
"An equivalent plot using the Toburen numbers n_e of Table II does not exhibit the straight lines obtained with the present scaling."

8) We appreciate the suggestion of the Z-scaling at intermediate energies. We are currently investigating its implementation. We refer to it in the manuscript by including the sentences:
"At intermediate impact energies, the Z^2 rule no longer holds, and other scalings can be considered in this region. For example, the molecular cross section and ion impact energy can be reduced with the projectile charge Z, as suggested in [33, 34]."

@@

Referee 2:

We decided to include in the introduction a comment about different previous work combining atomic cross sections to obtain molecular ones by Galassi et al. and Ludde et al.
"Different approaches have been proposed to deal with the ionization of molecular targets within the independent atom model. For example, Galassi et al. [2] obtain molecular cross sections by combining CDW-EIS atomic ones based on the population of the molecular orbitals. More recently, Lüdde et al. [3, 4] propose a combination of atomic cross sections with geometrical screening corrections."

### 2019-12-02 ionmol_stoi.tex
include Z scaling work

%\documentclass[pra,showpacs,twocolumn]{revtex4}
%\documentclass[pra,showpacs]{revtex4}
%\documentclass[preprint,showpacs]{revtex4}
\documentclass[10pt,showpacs,twocolumn]{revtex4}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{gensymb}
\usepackage{float}
\usepackage{MnSymbol}
\usepackage{gensymb}

@@

\title[Ionization of biological molecules by multicharged 
ions]{Ionization of biological molecules by multicharged ions using 
the stoichiometric model}
\author{A. M. P. Mendez, C. C. Montanari, J. E. Miraglia}
\address{Instituto de Astronom\'{\i}a y F\'{\i}sica del Espacio 
(CONICET--UBA), \\ Buenos Aires, Argentina.}

@@

\begin{abstract}
In the present work, we investigate the ionization of molecules of 
biological interest by the impact of multicharged ions in the 
intermediate to high energy range. We performed full non--perturbative 
distorted--wave calculations (CDW) for thirty--six collisional systems 
composed by six atomic targets: H, C, N, O, F and S --which are the 
constituents of most of the DNA and biological molecules-- and six 
charged projectiles (antiprotons, H, He, B, C, and O). On account of 
the radiation damage caused by secondary electrons, we inspect the 
energy and angular distributions of the emitted electrons from the 
atomic targets. We examine seventeen molecules: DNA and RNA bases, 
DNA backbone, pyrimidines, tetrahydrofuran (THF), and C$_n$H$_n$ 
compounds. We show that the simple stoichiometric model (SSM), which 
approximates the molecular ionization cross sections as a linear 
combination of the atomic ones, gives reasonably good results for 
complex molecules. We also inspect the extensively used Toburen scaling 
of the total ionization cross sections of molecules with the number of 
weakly bound electrons. Based on the atomic CDW results, we 
propose new active electron numbers, which leads to a more universal 
scaling for all the targets and ions studied here in the intermediate 
to the high energy region. The new scaling describes well the available 
experimental data for proton impact, including small molecules. We 
perform full molecular calculations for five nucleobases and test a 
modified stoichiometric formula based on the Mulliken charge of the 
composite atoms. The difference introduced by the new stoichiometric 
formula is less than 3\%, which indicates the reliability of the SSM to 
deal with this type of molecules. The results of the extensive 
ion--target examination included in the present study allow us to assert 
that the SSM and the CDW--based scaling will be useful tools in this area.
\end{abstract}

%\keywords{magnetic moment, solar neutrinos, astrophysics}
\pacs{34.50Gb, 34.80Gs, 34.80Dp}

@@

Different approaches have been proposed to deal with the ionization of 
molecular targets within the independent atom model. For example, 
Galassi \textit{et al.} \cite{galassi2000} obtains molecular cross 
sections by combining CDW-EIS atomic ones based on the population of 
the molecular orbitals. Much more recently, L\"udde \textit{et al.} 
\cite{ludde2016,ludde2018} 
propose a combination of atomic cross sections with geometrical screening
corrections.

The objective of this article is to deal with two aspects; first, we perform 
more appropriate calculations on the primary damage mechanism, which can 
replace the Born results. Second, we inspect and test a stoichiometric 
model to describe the ionization of molecular targets.

To overcome the first perturbative approximation limitations, and since 
the projectiles are multicharged ions, we resort to the Continuum 
Distorted Wave--Eikonal Initial State 
(CDW)~\cite{fainstein1988,miraglia2008,miraglia2009}, which includes higher 
perturbative corrections. Details on the CDW calculation are given in 
Section~\ref{sec:atoms}.
We start from the premise that the ionization 
process is the mechanism that deposits the most significant amount of 
primary energy. Moreover, the residual electrons from the ionization 
are known to be a source of significant local biological damage. 
The secondary electrons are included in Monte Carlo simulations, 
and hence their behavior requires further investigation.
In Section~\ref{subsec:meanener} and \ref{subsec:meanang}, we calculate
the mean energy and angular distributions of the ejected electrons. 
Surprisingly, we found a substantial dependence on the projectile charge, 
which is unexpected in the first Born approximation.

@@

In a given biological medium, direct ionization by ion impact accounts 
for just a fraction of the overall damage. Secondary electrons, as well 
as recoil target ions, also contribute substantially to the total damage. 
We can consider the single differential cross section of the shell 
$nl$ of the atom $\alpha$, $d\sigma_{\alpha nl}/dE$, to be a function 
of the ejected electron energy $E$ as a simple distribution 
function~\cite{surdutovic2018}. Then, we can define the mean value 
$\overline{E}_{\alpha}$ as in Ref.~\cite{abril2015},
\begin{eqnarray}
\overline{E}_{\alpha} &=&\frac{\langle E_{\alpha}\rangle}{\langle
1\rangle}=\frac{1}{\sigma_{\alpha}}\sum\limits_{nl}\int dE\,E
\frac{d\sigma_{\alpha,nl}}{dE}\,,  
\label{40} \\
\langle 1\rangle &=&\sigma_{\alpha}=\sum\limits_{nl}\int dE\,
\frac{d\sigma_{\alpha,nl}}{dE}\,,  
\label{50}
\end{eqnarray}
where $\Sigma_{nl}$ takes into account the sum of the different 
sub--shell contributions of the element $\alpha$.

@@

The mean emitted electron angles $\overline{\theta}_{\alpha}$ for the 
six atoms and six ions of interest are shown in Fig.~\ref{fig:emittedang}.
An significant dependence of $\overline{\theta}_{\alpha}$ with $Z$ is 
noticed for all the cases. Once again, this effect could not be 
observed in the first Born approximation (dotted line).

It is a general belief~\cite{Rudd1992} that the 
angular dispersion of emitted electrons is nearly isotropic. This 
behavior is thought to be caused by the insignificant angular anisotropy 
of sub--50--eV yield. 
A typical value for the ejection angle considered in the literature is 
$\overline{\theta}_{\alpha}\sim 70\degree$~\cite{surdutovic2018}, and 
it is quite correct in the range of validity of the first Born 
approximation for any target. However, when a distorted wave approximation 
is used, $\overline{\theta}_{\alpha}$ decreases substantially with $Z$ 
in the intermediate energy region, as shown in Fig.~\ref{fig:emittedang}.

To illustrate these effects, consider the impact of C$^{6+}$ impact on oxygen.
At an impact energy of 500~keV, the first Born approximation predicts
emitted electrons with mean energies of 46.7 eV and mean angles of 78\degree, 
while the CDW gives 62.5 eV and 60\degree. 
These results imply depper penetration closer to the direction of the ion.
We can attribute this forward direction correction to the capture to 
the continuum effect; the higher the charge $Z$, the smaller 
$\overline{\theta}$ will be. Of course, this effect only holds at 
intermediate energies and not at high impact energies.

@@

In Fig.~\ref{fig:crossDNA_1}, we report the reduced total ionization cross 
sections $\sigma_M/Z^2$ by the impact of multicharged ions for adenine,
cytosine, 
guanine, and thymine, by combining the SSM and CDW results in 
Eq.~\ref{eq:sumion}. For adenine, the agreement with the experimental 
data available for proton impact~\cite{iriki2011} is excellent. To the 
best of our knowledge, there are no experimental data on ion--collision 
ionization for the rest of the molecules. We have also included in this 
figure electron impact measurements~\cite{rahman2016} with the 
corresponding equivelocity conversion for electron incident energies 
higher than 300~eV. In this region, the proton and electron cross 
section should converge. Although the electron impact measurements are 
above our findings for all the molecular targets, it is worth stating 
that our results agree very well with other electron impact theoretical 
predictions~\cite{mozejko2003,tan2018}.

At intermediate impact energies, the $Z^2$ rule no longer holds and other 
scalings can be considered in this region. 
In Fig.~\ref{fig:Zcaled_crossDNA_1}, we plotted the reduced ionization 
cross section $\sigma_M/Z$ as a function of the reduced ion impact energy
$E/Z$ as suggested by~\cite{janev1980,dubois2013}.
This scaling may be a good alternative to deal with highly charged
ions at intermediate energies in a coherent way.

@@

@@

\begin{figure*}[t!]
\centering
\includegraphics[width=0.8\textwidth]{figuras/Zscaling_adn1.eps}
\caption{Reduced CDW ionization cross section $\sigma_{M}/Z$ as a function 
of reduced ion impact energy $E/Z$. 
Experiments: \mbox{\Large$\circ$}~\cite{iriki2011} 
for proton impact and $\square$~\cite{rahman2016} for electron impact 
with equivelocity conversion.}
\label{fig:Zcaled_crossDNA_1}
\end{figure*}

@@

By using Eq.~(\ref{eq:scalingCDW}), we define new active electron 
numbers $n_e'$ for some molecules of interest in Table~\ref{nn}. These 
values are different from the ones proposed by Toburen and used by 
other authors~\cite{itoh2013}. Moreover, an alternative way of showing 
the scaling can be attained by plotting the ionization cross sections 
of molecules as a function of the number of active electrons from 
Table~\ref{nn}. Our findings are displayed in Fig.~\ref{fig:recta} for 
impact energies 0.5, 1, and 2 MeV. As can be noted, the computed CDW 
ionization cross sections for all the molecules show a linear 
dependence with the number of electrons $n_e'$ from Table~\ref{nn}.
We obtain similar results, even for $E=10$~MeV. An 
equivalent plot using the Toburen numbers $n_e$ of Table~\ref{nn} does not 
exhibit the straight lines obtained with the present scaling. The 
comparison with the 
experimental data available shows overall good agreement, for the 
smallest molecules, H$_2$, H$_2$O, and CH$_4$, up to the most complex 
ones, like adenine. For electron impact data, the experimental data was 
interpolated between close neighbors.

While finishing the present work, we became aware of an accepted manuscript by 
Luedde~{\it et al.}~\cite{luedde2019} on total ionization of biological
molecules by proton impact, using an independent--atom--model pixel counting
method. The authors also raised a scaling with $\nu_{\alpha}=4$ 
for C, N, and O, but $\nu_{\alpha}=6$ for P. The agreement with this 
independent method for proton impact reinforces our multicharged--ion findings.

@@

To test the range of validity of the SSM, we performed {\it ab initio}
molecular calculation of five nucleobases by employing the {\sc gamess} 
code. The geometry optimization and single point energy calculations 
were performed implementing the restricted Hartree--Fock method and the 
3-21G basis set.

@@

By implementing Eq.~(\ref{eq:newstoi}), it is possible to determine a 
new stoichiometric formula (last column of Table~\ref{tab:newstoi}). 
Now, instead of having an integer number of atoms $n_{\alpha}$, we obtain 
a fractional number $n_{\alpha}^{\prime}$. New molecular cross sections 
$\sigma^{\prime}_{M}=\sum_{\alpha}n_{\alpha}'\sigma_{\alpha}$ can be 
computed considering 
such values. Relative errors for the ionization cross sections were 
computed for the DNA bases from Table~\ref{tab:newstoi}. The differences 
obtained were less than 3\%, which indicates that the modified 
SSM is a quite robust model to handle these type of molecules within 
the range error expected for this model.

@@

In this work, we have dealt with the calculation of ionization cross 
sections of seventeen biological molecules containing H, C, N, O, P, and 
S by the impact of antiprotons, H$^{+}$, He$^{2+}$, Be$^{4+}$, C$^{6+}$, 
and O$^{8+}$. To that end, we have employed the full CDW method and the 
simple stoichiometric model. 
The mean energy and angle of the emitted electrons, of importance in 
post--collisional radiation damage,  has also been calculated, showing 
a clear dependence with the ion charge $Z$. For a given target as $Z$ 
increases, $\overline{E}_{\alpha}$
increases, but $\overline{\theta}_{\alpha}$ decreases, showing a clear 
tendency to the forward direction. At impact
energies greater than 2 MeV/amu, these values converge to the Born
approximation, which embodies the simple $Z^{2}$ law.

Total ionization cross sections for adenine, cytosine, thymine, guanine, 
uracil, DNA backbone, pyrimidine, and THF are presented and compared 
with the scarcely available experiments. We explore the rule of 
Toburen, which scales all the molecular ionization cross section 
normalizing with a certain number of weakly bound or valence electrons. 
We found that the ionization cross 
sections scales much better when normalizing with the number of active 
electrons in the collision obtained from the CDW results for atoms. 
This new scaling was tested with good results for the six 
projectiles and seventeen molecules studied here. The comparison with 
the experimental data reinforce these results. Furthermore, we also tested
the scaling by including experimental data of ionization of H$_2$, 
water, methane, and ammonia by proton impact showing good agreement at
intermediate to high energies.

@@

\bibitem{Mohamad2017}
O. Mohamad, B. J. Sishc, J. Saha, A. Pompos, A. Rahimi, M. D. Story, A. J. Davis, D. N. Kim, 
Carbon Ion Radiotherapy: A Review of Clinical Experiences and Preclinical Research, with an Emphasis on DNA Damage/Repair. 
Cancers \textbf{9}, 66 (2017).

@@

\bibitem{miraglia2008} 
J. E. Miraglia and M. S. Gravielle. 
Ionization of the He, Ne, Ar, Kr, and Xe isoelectronic series by proton impact. 
Phys Rev A \textbf{78}, 052705 (2008)

\bibitem{miraglia2009} 
J. E. Miraglia, Ionization of He, Ne, Ar, Kr, and Xe
by proton impact: Single differential distributions. 
Phys. Rev. A \textbf{79}, 022708 (2009).

\bibitem{toburen1975} 
W. E. Wilson and L. H. Toburen. 
Electron emission from proton --hydrocarbon-molecule collisions at 0.3--2.0 MeV. 
Phys. Rev. A \textbf{11}, 1303 (1975).

\bibitem{toburen1976} 
D. J. Lynch, L. H. Toburen, and W. E. Wilson. 
Electron emission from methane, ammonia, monomethylamine, and dimethylamine by 0.25
to 2.0 MeV protons. 
J. Chem. Phys. \textbf{64}, 2616 (1976).

@@

\bibitem{mendez2016} 
A.M.P. Mendez, D.M. Mitnik, and J.E. Miraglia.
Depurated inversion method for orbital-specific exchange potentials. 
Int. J. Quantum Chem. \textbf{24} ,116 (2016).

\bibitem{mendez2018} 
A.M.P. Mendez, D.M. Mitnik, and J.E. Miraglia. 
Local Effective Hartree--Fock Potentials Obtained by the Depurated Inversion Method,
Adv. Quant. Chem. \textbf{76}, 117--132 (2018).

\bibitem{montanari2017} 
C. C. Montanari, J. E. Miraglia,
Ionization probabilities of Ne, Ar, Kr, and Xe by proton impact for different initial states and impact energies. 
Nucl. Instr. Meth. Phys. Res. B \textbf{407}, 236--243 (2017).

\bibitem{miraglia2019} 
J. E. Miraglia,
Shell-to-shell ionization cross sections of antiprotons, H$^{+}$, He$^{2+},$ Be$^{4+},$ C$^{6+}$ and 
O$^{8+}$ on H, C, N, O, P, and S atoms (to be published).

\bibitem{surdutovic2018} 
E. Surdutovich and A. V. Solov'yov, 
Multiscale approach to the physics of radiation damage with ions. 
arXiv:1312.0897v, (2013)

\bibitem{abril2015} 
P. de Vera1, I. Abril, R. Garcia-Molina and
A. V. Solov'yov,
Ionization of biomolecular targets by ion impact: input data for radiobiological applications. 
Journal of Physics: Conference Series \textbf{438}, 012015 (2013).

\bibitem{Rudd1992} 
M. E. Rudd, Y.-K. Kim,, D. H. Madison and T. J. Gay.
Electron production in proton collisions with atoms and molecules: energy distributions. 
Rev. Mod. Phys. \textbf{64}, 44--490 (1992).

@@

\bibitem{mozejko2003}
P. Mozejko and L. Sanche, 
Cross section calculations for electron scattering from DNA and RNA bases.
Radiat Environ. Biophys \textbf{42}, 201 (2003).

\bibitem{tan2018}
H. Q. Tan, Z. Mi, and A. A. Bettiol, 
Simple and universal model for electron-impact ionization of complex biomolecules, 
Phys. Rev. E \textbf{97}, 032403 (2018)

>>>>

\documentclass[10pt,showpacs,showkeys,twocolumn]{revtex4}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{gensymb}
\usepackage{float}
\usepackage{MnSymbol}
\usepackage{gensymb}

@@

\title[Ionization of biological molecules by multicharged 
ions]{Ionization of biological molecules by multicharged ions using 
the stoichiometric model}
\author{A. M. P. Mendez, C. C. Montanari, J. E. Miraglia}
\affiliation{Instituto de Astronom\'{\i}a y F\'{\i}sica del Espacio 
(CONICET--UBA), \\ Buenos Aires, Argentina.}

@@

\begin{abstract}
In the present work, we investigate the ionization of molecules of 
biological interest by the impact of multicharged ions in the 
intermediate to high energy range. We performed full non--perturbative 
distorted--wave calculations (CDW) for thirty--six collisional systems 
composed by six atomic targets: H, C, N, O, F, and S --which are the 
constituents of most of the DNA and biological molecules-- and six 
charged projectiles (antiprotons, H, He, B, C, and O). On account of 
the radiation damage caused by secondary electrons, we inspect the 
energy and angular distributions of the emitted electrons from the 
atomic targets. We examine seventeen molecules: DNA and RNA bases, 
DNA backbone, pyrimidines, tetrahydrofuran (THF), and C$_n$H$_n$ 
compounds. We show that the simple stoichiometric model (SSM), which 
approximates the molecular ionization cross sections as a linear 
combination of the atomic ones, gives reasonably good results for 
complex molecules. We also inspect the extensively used Toburen scaling 
of the total ionization cross sections of molecules with the number of 
weakly bound electrons. Based on the atomic CDW results, we 
propose new active electron numbers, which leads to a better universal 
scaling for all the targets and ions studied here in the intermediate 
to the high energy region. The new scaling describes well the available 
experimental data for proton impact, including small molecules. We 
perform full molecular calculations for five nucleobases and test a 
modified stoichiometric formula based on the Mulliken charge of the 
composite atoms. The difference introduced by the new stoichiometric 
formula is less than 3\%, which indicates the reliability of the SSM to 
deal with this type of molecules. The results of the extensive 
ion--target examination included in the present study allow us to assert 
that the SSM and the CDW--based scaling will be useful tools in this area.
\end{abstract}

\keywords{DNA bases, DNA ionization, stoichiometric model, 
multicharged ions, biological molecules, ionization of molecules}
\pacs{34.50Gb, 34.80Gs, 34.80Dp}

@@

Different approaches have been proposed to deal with the ionization of 
molecular targets within the independent atom model. For example, 
Galassi \textit{et al.} \cite{galassi2000} obtain molecular cross 
sections by combining CDW-EIS atomic ones based on the population of 
the molecular orbitals. More recently, L\"udde \textit{et al.} 
\cite{ludde2016,ludde2018} 
propose a combination of atomic cross sections with geometrical screening
corrections.

The objective of this article is to face with two aspects of the ionization
of biological molecules; first, we perform 
more appropriate calculations on the primary damage mechanism, which can 
replace the Born results. Second, we inspect and test a stoichiometric 
model to describe the ionization of molecular targets.

To overcome the limitations of first order perturbative approximations, 
and since the projectiles are multicharged ions, we resort to the 
Continuum Distorted Wave--Eikonal Initial State 
(CDW)~\cite{galassi2000,fainstein1988,miraglia2008,miraglia2009}, which 
includes higher perturbative corrections. Details on the CDW calculation 
are given in Section~\ref{sec:atoms}.
We start from the premise that the ionization 
process is the mechanism that deposits the most significant amount of 
primary energy. Moreover, the residual electrons from the ionization 
are known to be a source of significant local biological damage~\cite{Denifl2011}. 
The secondary electrons are included in Monte Carlo simulations, 
and hence their behavior requires further investigation.
In Section~\ref{subsec:meanener} and \ref{subsec:meanang}, we calculate
the mean energy and angular distributions of the ejected electrons. 
Surprisingly, we found a substantial dependence on the projectile charge, 
which is unexpected in the first Born approximation.

@@

In a given biological medium, direct ionization by ion impact accounts 
for just a fraction of the overall damage. Secondary electrons, as well 
as recoil target ions, also contribute substantially to the total damage~\cite{Denifl2011}. 
We can consider the single differential cross section of the shell 
$nl$ of the atom $\alpha$, $d\sigma_{\alpha nl}/dE$, to be a function 
of the ejected electron energy $E$ as a simple distribution 
function~\cite{surdutovic2018}. Then, we can define the mean value 
$\overline{E}_{\alpha}$ as in Ref.~\cite{abril2015},
\begin{eqnarray}
\overline{E}_{\alpha} &=&\frac{\langle E_{\alpha}\rangle}{\langle
1\rangle}=\frac{1}{\sigma_{\alpha}}\sum\limits_{nl}\int dE\,E
\frac{d\sigma_{\alpha,nl}}{dE}\,,  
\label{40} \\
\langle 1\rangle &=&\sigma_{\alpha}=\sum\limits_{nl}\int dE\,
\frac{d\sigma_{\alpha,nl}}{dE}\,,  
\label{50}
\end{eqnarray}
where $\Sigma_{nl}$ takes into account the sum of the different 
sub--shell contributions of the element $\alpha$.

@@

The mean emitted electron angles $\overline{\theta}_{\alpha}$ for the 
six atoms and six ions of interest are shown in Fig.~\ref{fig:emittedang}.
A significant dependence of $\overline{\theta}_{\alpha}$ with $Z$ is 
noticed for all the cases. Once again, this effect could not be 
observed in the first Born approximation (dotted line).

For low energy electron emission, the angular dispersion is nearly 
isotropic~\cite{Rudd1992}.
A typical value for the ejection angle considered in the literature is 
$\overline{\theta}_{\alpha}\sim 70\degree$~\cite{surdutovic2018}, and 
it is quite correct in the range of validity of the first Born 
approximation for any target. However, when a distorted wave approximation 
is used, $\overline{\theta}_{\alpha}$ decreases substantially with $Z$ 
in the intermediate energy region, as shown in Fig.~\ref{fig:emittedang}.
The higher the charge $Z$, the smaller 
$\overline{\theta}$ will be. Of course, this effect only holds at 
intermediate energies and not at high impact energies.

To illustrate this feature, consider the impact of 500~keV C$^{6+}$ on 
oxygen. The first Born approximation predicts emitted electrons with 
mean energies of 46.7 eV and mean angles of 78\degree, while the CDW 
gives 62.5 eV and 60\degree. 
These results imply deeper penetration of the secondary electrons with 
an orientation closer to the direction of the ion. 
We can attribute this forward direction correction to the capture to 
the continuum effect.

@@

In Fig.~\ref{fig:crossDNA_1}, we report the reduced total ionization cross 
sections $\sigma_M/Z^2$ for adenine, cytosine, guanine, and thymine 
by the impact of multicharged ions obtained combining the SSM given by 
Eq.~(\ref{eq:sumion}) and the CDW results. For adenine, the agreement 
with the experimental 
data available for proton impact~\cite{iriki2011} is excellent. To the 
best of our knowledge, there are no experimental data on ion--collision 
ionization for the rest of the molecules. We have also included in this 
figure electron impact measurements~\cite{rahman2016} with the 
corresponding equivelocity conversion for electron incident energies 
higher than 300~eV. In this region, the proton and electron cross 
section should converge. Although the electron impact measurements are 
above our findings for all the molecular targets, it is worth stating 
that our results agree very well with other electron impact theoretical 
predictions~\cite{mozejko2003,tan2018}.

@@

At intermediate impact energies, the $Z^2$ rule no longer holds, and 
other scalings can be considered in this region. For example, the 
molecular cross section and ion impact energy can be reduced with the
projectile charge $Z$, as suggested in in~\cite{janev1980,dubois2013}.

@@

@@

By using Eq.~(\ref{eq:scalingCDW}), we define new active electron 
numbers $n_e'$ for molecules. In Table~\ref{nn}, we display the present 
$n_e'$ values and $n_e$ ones by Toburen obtained from Eq.~(\ref{eq:nelec}).
Our values are different from the ones proposed by Toburen and used by 
other authors~\cite{itoh2013}, mainly due to the differences in the 
active electron numbers of oxygen. An alternative way of testing 
the present scaling can be attained by plotting the ionization cross sections 
of molecules as a function of the $n_e'$ from Table~\ref{nn}. Our 
findings are displayed in Fig.~\ref{fig:recta} for 
impact energies 0.5, 1, and 2 MeV. As can be noted, the computed CDW 
ionization cross sections for all the molecules show a linear 
dependence with the number of electrons $n_e'$ from Table~\ref{nn}.
We obtain similar results, even for $E=10$~MeV. The comparison with the 
experimental data available shows overall good agreement, for the 
smallest molecules, H$_2$, H$_2$O, and CH$_4$, up to the most complex 
ones, like adenine. For electron impact data, the experimental data was 
interpolated between close neighbors.
It is worth mentioning that an equivalent plot using the Toburen numbers 
$n_e$ does not exhibit the straight lines obtained with the present scaling.

While finishing the present work, we became aware of an accepted manuscript by 
L\"udde~{\it et al.}~\cite{ludde2019} on total ionization of biological
molecules by proton impact, using the independent--atom--model pixel counting
method~\cite{ludde2016,ludde2018}. The authors also raised a scaling 
with $\nu_{\alpha}=4$ 
for C, N, and O, but $\nu_{\alpha}=6$ for P. The agreement with this 
independent method for proton impact reinforces our multicharged--ion findings.

@@

Finally, to test the range of validity of the SSM, we performed {\it ab initio}
molecular calculation of five nucleobases by employing the {\sc gamess} 
code. The geometry optimization and single point energy calculations 
were performed implementing the restricted Hartree--Fock method and the 
3-21G basis set.

@@

By implementing Eq.~(\ref{eq:newstoi}), it is possible to determine a 
new stoichiometric formula (last column of Table~\ref{tab:newstoi}). 
Now, instead of having an integer number of atoms $n_{\alpha}$, we obtain 
a fractional number $n_{\alpha}^{\prime}$. New molecular cross sections 
$\sigma^{\prime}_{M}=\sum_{\alpha}n_{\alpha}'\sigma_{\alpha}$ can be 
computed considering 
such values. Relative errors for the ionization cross sections were 
computed for the DNA bases from Table~\ref{tab:newstoi}. The differences 
obtained were less than 3\%, which indicates that the  
SSM is a quite robust model to handle these type of molecules within 
the range error expected for this model.

@@

In this work, we have dealt with the calculation of ionization cross 
sections of seventeen biological molecules containing H, C, N, O, P, and 
S by the impact of antiprotons, H$^{+}$, He$^{2+}$, Be$^{4+}$, C$^{6+}$, 
and O$^{8+}$. To that end, we have employed the full CDW method and the 
simple stoichiometric model. 
The mean energy and angle of the emitted electrons, of importance in 
post--collisional radiation damage,  has also been calculated. Our 
findings show a clear dependence with the ion charge $Z$. For a given 
target as $Z$ increases, $\overline{E}_{\alpha}$ also increases, but 
$\overline{\theta}_{\alpha}$ decreases, showing a clear tendency to the 
forward direction. At impact energies greater than 2 MeV/amu, these 
values converge to the Born approximation, which embodies the simple 
$Z^{2}$ law.

Total ionization cross sections for adenine, cytosine, thymine, guanine, 
uracil, DNA backbone, pyrimidine, and THF are presented and compared 
with the scarcely available experiments. We explored the rule of 
Toburen, which scales all the molecular ionization cross section 
normalizing with a certain number of weakly bound or valence electrons. 
We found that the ionization cross 
sections scales much better when normalizing with the number of active 
electrons in the collision obtained from the CDW results for atoms. 
This new scaling was tested with good results for the six 
projectiles and seventeen molecules studied here. The comparison with 
the experimental data reinforce our findings. Furthermore, we tested
the scaling by including experimental data of ionization of H$_2$, 
water, methane, and ammonia by proton impact showing good agreement at
intermediate to high energies.

@@

\bibitem{Mohamad2017}
O. Mohamad, B. J. Sishc, J. Saha, A. Pompos, A. Rahimi, M. D. Story, A. J. Davis, D. N. Kim, 
%Carbon Ion Radiotherapy: A Review of Clinical Experiences and Preclinical Research, with an Emphasis on DNA Damage/Repair. 
Cancers \textbf{9}, 66 (2017).

@@

\bibitem{miraglia2008} 
J. E. Miraglia and M. S. Gravielle. 
%Ionization of the He, Ne, Ar, Kr, and Xe isoelectronic series by proton impact. 
Phys Rev A \textbf{78}, 052705 (2008)

\bibitem{miraglia2009} 
J. E. Miraglia, 
%Ionization of He, Ne, Ar, Kr, and Xe by proton impact: Single differential distributions. 
Phys. Rev. A \textbf{79}, 022708 (2009).

\bibitem{Denifl2011}
Denifl S., Märk T.D., Scheier P. 
The Role of Secondary Electrons in Radiation Damage. 
In Radiation Damage in Biomolecular Systems. Biological and Medical Physics, Biomedical Engineering. 
Eds: García Gómez-Tejedor G., Fuss M. 
Springer, Dordrecht (2012)

\bibitem{toburen1975} 
W. E. Wilson and L. H. Toburen. 
%Electron emission from proton --hydrocarbon-molecule collisions at 0.3--2.0 MeV. 
Phys. Rev. A \textbf{11}, 1303 (1975).

\bibitem{toburen1976} 
D. J. Lynch, L. H. Toburen, and W. E. Wilson. 
%Electron emission from methane, ammonia, monomethylamine, and dimethylamine by 0.25 to 2.0 MeV protons. 
J. Chem. Phys. \textbf{64}, 2616 (1976).

@@

\bibitem{mendez2016} 
A.M.P. Mendez, D.M. Mitnik, and J.E. Miraglia.
%Depurated inversion method for orbital-specific exchange potentials. 
Int. J. Quantum Chem. \textbf{24} ,116 (2016).

\bibitem{mendez2018} 
A.M.P. Mendez, D.M. Mitnik, and J.E. Miraglia. 
%Local Effective Hartree--Fock Potentials Obtained by the Depurated Inversion Method,
Adv. Quant. Chem. \textbf{76}, 117--132 (2018).

\bibitem{montanari2017} 
C. C. Montanari, J. E. Miraglia,
%Ionization probabilities of Ne, Ar, Kr, and Xe by proton impact for different initial states and impact energies. 
Nucl. Instr. Meth. Phys. Res. B \textbf{407}, 236--243 (2017).

\bibitem{miraglia2019} 
J. E. Miraglia,
%Shell-to-shell ionization cross sections of antiprotons, H$^{+}$, He$^{2+},$ Be$^{4+},$ C$^{6+}$ and 
O$^{8+}$ on H, C, N, O, P, and S atoms (to be published).

\bibitem{surdutovic2018} 
E. Surdutovich and A. V. Solov'yov, 
%Multiscale approach to the physics of radiation damage with ions. 
arXiv:1312.0897v, (2013)

\bibitem{abril2015} 
P. de Vera, I. Abril, R. Garcia-Molina and A. V. Solov'yov,
%Ionization of biomolecular targets by ion impact: input data for radiobiological applications. 
Journal of Physics: Conference Series \textbf{438}, 012015 (2013).

\bibitem{Rudd1992} 
M. E. Rudd, Y.-K. Kim,, D. H. Madison and T. J. Gay.
%Electron production in proton collisions with atoms and molecules: energy distributions. 
Rev. Mod. Phys. \textbf{64}, 441--490 (1992).

@@

\bibitem{mozejko2003}
P. Mozejko and L. Sanche, 
%Cross section calculations for electron scattering from DNA and RNA bases.
Radiat Environ. Biophys \textbf{42}, 201 (2003).

\bibitem{tan2018}
H. Q. Tan, Z. Mi, and A. A. Bettiol, 
%Simple and universal model for electron-impact ionization of complex biomolecules, 
Phys. Rev. E \textbf{97}, 032403 (2018)

\bibitem{itoh2013} 
A. Itoh, Y. Iriki, M. Imai, C. Champion, and R. D. Rivarola, 
%Cross sections for ionization of uracil by MeV-energy-proton impact, 
Phys. Rev. A \textbf{88}, 052711 (2013).

\bibitem{agnihotri2012}
A. N. Agnihotri, S. Kasthurirangan, S. Nandi, A.
Kumar, M. E. Galassi, R. D. Rivarola, O. Foj\'{o}n, C. Champion, J. Hanssen,
H. Lekadir, P. F. Weck, and L. C. Tribedi. 
%Ionization of uracil in collisions with highly charged carbon and oxygen ions of energy 100 keV to 78 MeV. 
Phys. Rev. A \textbf{85}, 032711 (2012).

\bibitem{agnihotri2013}
A N Agnihotri, S Kasthurirangan, S Nandi, A Kumar, C Champion,, H Lekadir, 
J Hanssen, P FWeck, M E Galassi, R D Rivarola, O Fojon and L C Tribedi, 
%Absolute total ionization cross sections of uracil (C$_4$H$_4$N$_2$O$_2$) in collisions with MeV energy highly charged carbon, oxygen and fluorine ions
J. Phys. B \textbf{46}, 185201 (2013).
